---
title: "Module 10"
author: "Amy Scott"
date: "10/3/2017"
output: html_document
---

##Classical hypothesis testing
this is only for frequentist data, normal/quassian

T-test, one-tailed and two-tailed

*this test only applies to normal data*

 Ho= TINDB the two years
 
 Ha= 2016>2015
```{r}
library(curl)
v <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall17/vervet-weights.csv")
vervets <- read.csv(v, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(vervets)
t.test(x = vervets$weight, mu=4.9, alternative = "greater")

mu <- 4.9 # known mean
x <- vervets$weight
m <- mean(x)
s <- sd(x)
n <- length(x)
sem <- s/sqrt(n)

lower <- m - qt(1 - 0.05/2, df = n - 1) * sem
upper <- m + qt(1 - 0.05/2, df = n - 1) * sem
ci <- c(lower, upper)
ci  # by hand

```

Now we are comparing two samples to see if they are the same or different

*this test only applies to normal data*

For the most generic case, where the two samples are independent and we cannot assume the variances of the two samples are equal, we can do what is called Welch’s t test
```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall17/colobus-weights.csv")
colobus<- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)

#t <- t.test(x = x, y = y, mu = 0, alternative = "two.sided")
m <- colobus$weight[colobus$sex == "male"]
f <- colobus$weight[colobus$sex == "female"]
#mu <- 0  # you could leave this out... the default argument value is 0
#t.test( one group, 2nd group, alternaive= "two.sided")
t.test(m, f, alternative= "two.sided")
```

For the case, where the two samples are independent and we can assume the variances of the two samples are equal, what we do is:

```{r}
t <- t.test(x = x, y = y, mu = 0, var.equal = TRUE, alternative = "two.sided")
#t.test(one group, two group, mu=0, var.equal=TRUE, alternative= "two.sided" )
```

Paired Samples

```{r}
f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN597_Fall17/iqs.csv")
IQ <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)


x <- IQ$IQ.before - IQ$IQ.after

t.test(x, df = n - 1, alternative = "two.sided")
# OR you can use paired="TRUE" in the , but i dont remember how to do that, so look up. I did this when i looked at the difference in distance when a male was and was not there.

```
Chi-squares: comparing differences in proportions


