---
title: "Module 9 Introduction to Statistical Inference"
author: "Amy Scott"
date: "10/3/2017"
output: html_document
---

```{r}
library(curl)
n <- 1000
mu <- 3.5
sigma <- 4
v <- rnorm(n, mu, sigma)
s <- sample(v, size = 30, replace = FALSE)
m <- mean(s)
m
sd <- sd(s)
sd
sem <- sd(s)/sqrt(length(s))
sem
lower <- m - qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
upper <- m + qnorm(1 - 0.05/2) * sem  # (1-alpha)/2 each in the upper and lower tails of the distribution
ci <- c(lower, upper)
ci
```
Central Limit Theorem says that as you add data, the arithmatic mean will form a normal distribution, that is that the deviations of the measurements from the mean will also have a normal distribution. 
```{r}
lambda <- 14 # mean for a poisson distribution
n <- 10
pop_se <- sqrt(lambda/n)  # the estimated SE --have to use Standard error
pop_se
x <- NULL
for (i in 1:1000) {
    x[i] <- mean(rpois(n = n, lambda = lambda))
}
hist(x, breaks = seq(from = lambda - 4 * sqrt(lambda)/sqrt(n), to = lambda + 
    4 * sqrt(lambda)/sqrt(n), length.out = 20), probability = TRUE)
# the histogram graph of the means is normal, even though its a poisson distribution
qqnorm(x) # its normal
qqline(x) # this one didn't work

curve(dnorm(x, 0, 1), -4, 4, ylim = c(0, 0.8))
z <- (x - lambda)/pop_se
hist(z, breaks = seq(from = -4, to = 4, length.out = 20), probability = TRUE, 
    add = TRUE)
```

##Challenge
Suppose a polling group in the Massachusetts is interested in the proportion of voting-age citizens in their state that already know they will vote for Elizabeth Warren in the upcoming November 6, 2018 midterm elections (don’t forget to vote!). The group obtains a yes or no answer from 1000 suitable randomly selected individuals. Of these individuals, 856 say they know they’ll vote for Senator Warren. How would we characterize the mean and variability associated with this proportion?
```{r}
# this is for binomial data
n <- 1000
x <- 856
phat <- x/n  # our estimate of pi
phat
n * phat # has to be >5; its an assumption it has to fulfill
n * (1 - phat) # has to be >5; its an assumption it has to fulfill
pop_se <- sqrt((phat) * (1 - phat)/n)


curve(dnorm(x, mean = phat, sd = pop_se), phat - 4 * pop_se, phat + 4 * pop_se)
upper <- phat + qnorm(0.975) * pop_se
lower <- phat - qnorm(0.975) * pop_se
ci <- c(lower, upper)
polygon(cbind(c(ci[1], seq(from = ci[1], to = ci[2], length.out = 1000), ci[2]), 
    c(0, dnorm(seq(from = ci[1], to = ci[2], length.out = 1000), mean = phat, 
        sd = pop_se), 0)), border = "black", col = "gray")
abline(v = ci)
abline(h = 0)
```
